---
title: "COVID-19"
author: "Mike"
date: "3/12/2020"
output: 
  html_document:
    toc: TRUE
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## COVID Data Exploration

```{r, echo=FALSE,include=FALSE}

library(tidyverse)
library(lubridate)
library(reshape2)
library(viridis)
library(ggthemes)
library(plotly)
library(maps)

#install.packages('maps', dependencies=TRUE, repos='http://cran.rstudio.com/')

data <- read_csv("time_series_covid_19_confirmed.csv")

data.save <- data

data <- data[,-c(1,3,4)]

data$`Country/Region` <- ifelse(grepl("Mainland China", data$`Country/Region`),'China',
                         ifelse(grepl("Viet Nam", data$`Country/Region`),'Vietnam',
                         ifelse(grepl("Taiwan*", data$`Country/Region`),'Taiwan', 
                         ifelse(grepl("Iran (Islamic Republic of)", data$`Country/Region`),'Iran',
                         ifelse(grepl("Russian Federation", data$`Country/Region`),'Russia',
                         ifelse(grepl("occupied Palestinian territory", data$`Country/Region`),'Palestine',
                         data$`Country/Region`))))))

# Sum is required for the two "yes" responses added above
data <- data %>%
          group_by(`Country/Region`) %>% 
          summarise_all(list(sum))

# Melt into long form
data <- melt(data)

# Rename columns
colnames(data)[colnames(data)=="variable"] <- "date"
colnames(data)[colnames(data)=="value"] <- "count"

```

### Date Parsing and Formatting

Including the code below as it's nearly boilerplate for date formatting. In some cases the strsplit will need to be adjusted, but I find it much more efficient to parse and reformat a date each time than try to accomodate different date formats across datasets. 

```{r}

# Date to char for parsing and conversion
data$date <- as.character(data$date)

# Parse the date
data$year<-sapply(data$date, function(x) as.numeric(strsplit(x,"/")[[1]][3]))
data$month<-sapply(data$date, function(x) as.numeric(strsplit(x,"/")[[1]][1]))
data$day<-sapply(data$date, function(x) as.numeric(strsplit(x,"/")[[1]][2]))

# Put us back in the year 2000
data$year <- (data$year + 2000)

# Reformat date into y-m-d 
data$date<-as.Date(paste0(data$year,'-',data$month,'-',data$day), format="%Y-%m-%d")

# Add some weekdays for good measure
data$weekday <- weekdays(as.Date(data$date))

# Put weekdays in order
data$weekday <- ordered(data$weekday, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))


# Set X axis limits ------------------------------------------------------------------------------- Must update with data refresh
time <- as.POSIXct(strptime(c("2020-01-22","2020-03-15"), format = "%y-%m-%d"))

```

```{r, echo=FALSE, include=FALSE}
p <- ggplot(data, aes(date, `Country/Region`)) + 
     geom_tile(aes(fill = count)) +
     scale_fill_viridis(name="Confirmed Cases", option ="A", limits = c(0, 10000), oob = scales::squish) +
     labs(title = "Global Confirmed COVID-19 Cases by Country", subtitle = "Count of cases")

```

## Time Series Heat Plot 

At present, with China as a large confirmed case outlier (over 80,000), I rescaled the confirmed cases using a limit of 10,000 to give greater visibility to lower and emerging counts. 

```{r, echo=FALSE, fig.height=20, fig.width=18}

plot(p)

```


```{r, echo=FALSE, include=FALSE}
# Because the most recent date has missing data, I'm using the next most recent date
# ----------------------------------------------------------------------------------------------- Must update with data refresh
top <- data %>% filter(date == '2020-03-14') %>% arrange(desc(count)) %>% top_n(20, count)

trend <- data %>% filter(`Country/Region` %in% top$`Country/Region`) %>%
                  group_by(date, count) %>%
                  ggplot(aes(x=as.POSIXct(date),y=count, color=`Country/Region`)) + 
                  geom_point(size=.5) + 
                  geom_line(alpha=.5) + 
                  scale_x_datetime(limits =time) + 
                  labs(title="Total COVID-19 Confirmed Case Count") +
                  xlab("Date") + 
                  ylab("Cases")
  

```

<br>
<br>

## Confirmed Cases Trend

Using ggplotly for an interactive plot, here are the Country/Region's in the top 20 for confirmed cases plotted since the pandemic started.

```{r, echo=FALSE, fig.height=8, fig.width=10}

ggplotly(trend)

```


```{r, echo=FALSE, include=FALSE}

# process the death dataset in the same way as the confirmed cases

death <- read_csv("time_series_covid_19_deaths.csv")

death.save <- death

death <- death[,-c(1,3,4)]

death$`Country/Region` <- ifelse(grepl("Mainland China", death$`Country/Region`),'China',
                         ifelse(grepl("Viet Nam", death$`Country/Region`),'Vietnam',
                         ifelse(grepl("Taiwan*", death$`Country/Region`),'Taiwan', 
                         ifelse(grepl("Iran (Islamic Republic of)", death$`Country/Region`),'Iran',
                         ifelse(grepl("Russian Federation", death$`Country/Region`),'Russia',
                         ifelse(grepl("occupied Palestinian territory", death$`Country/Region`),'Palestine',
                         death$`Country/Region`))))))

# Sum is required for the two "yes" responses added above
death <- death %>%
          group_by(`Country/Region`) %>% 
          summarise_all(list(sum))

# Melt into long form
death <- melt(death)

# Rename columns
colnames(death)[colnames(death)=="variable"] <- "date"
colnames(death)[colnames(death)=="value"] <- "count"

# Date to char for parsing and conversion
death$date <- as.character(death$date)

# Parse the date
death$year<-sapply(death$date, function(x) as.numeric(strsplit(x,"/")[[1]][3]))
death$month<-sapply(death$date, function(x) as.numeric(strsplit(x,"/")[[1]][1]))
death$day<-sapply(death$date, function(x) as.numeric(strsplit(x,"/")[[1]][2]))

# Put us back in the year 2000
death$year <- (death$year + 2000)

# Reformat date into y-m-d 
death$date<-as.Date(paste0(death$year,'-',death$month,'-',death$day), format="%Y-%m-%d")

# Add some weekdays for good measure
death$weekday <- weekdays(as.Date(death$date))

# Put weekdays in order
death$weekday <- ordered(death$weekday, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

#---------------------------------------------------------------------------------------------------------Must update with new data
top.death <- death %>% filter(date == '2020-03-14') %>% arrange(desc(count)) %>% top_n(20, count)

trend.death <- death %>% filter(`Country/Region` %in% top.death$`Country/Region`) %>%
                  group_by(date, count) %>%
                  ggplot(aes(x=as.POSIXct(date),y=count, color=`Country/Region`)) + 
                  geom_point(size=.5) + 
                  geom_line(alpha=.5) + 
                  scale_x_datetime(limits =time) + 
                  labs(title="Global COVID-19 Deaths") +
                  xlab("Date") + 
                  ylab("Deaths")

```

<br>
<br>

## COVID-19 Death Trend 

Time series plot of top 20 County/Region's with COVID-19 related deaths

```{r, echo=FALSE, fig.height=8, fig.width=10}

ggplotly(trend.death)

```

<br>
<br>

## COVID-19 Confirmed Cases and Deaths {.tabset .tabset-fade}

Graphs of relationship between confirmed cases and death for highly impacted countries. 

### China

Example code included, overlaying two area plots from seperate data frames. I wasn't able to identify a straightforward method to include a plot legend-- comment below if you have one!

```{r}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "China") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color='red', alpha=0.4, fill = 'red') +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "China") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color='black') +
                  labs(title="China Confirmed Cases and Deaths", 
                       subtitle = " Red is confirmed cases \n Black is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```

### US

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "US") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color='red', alpha=0.4, fill = 'red') +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "US") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color='black') +
                  labs(title="US Confirmed Cases and Deaths", 
                       subtitle = " Red is confirmed cases \n Black is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```

### Italy

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "Italy") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color='red', alpha=0.4, fill = 'red') +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "Italy") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color='black') +
                  labs(title="Italy Confirmed Cases and Deaths", 
                       subtitle = " Red is confirmed cases \n Black is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```


### Iran

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "Iran") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color='red', alpha=0.4, fill = 'red') +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "Iran") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color='black') +
                  labs(title="Iran Confirmed Cases and Deaths", 
                       subtitle = " Red is confirmed cases \n Black is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```

### South Korea

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "Korea, South") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color='red', alpha=0.4, fill = 'red') +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "Korea, South") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color='black') +
                  labs(title="South Korea Confirmed Cases and Deaths", 
                       subtitle = " Red is confirmed cases \n Black is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```


### Spain

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "Spain") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color='red', alpha=0.4, fill = 'red') +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "Spain") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color='black') +
                  labs(title="Spain Confirmed Cases and Deaths", 
                       subtitle = " Red is confirmed cases \n Black is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```
```{r, echo=FALSE, include=FALSE}

#-----------------------------------------------------------------------------------------------Must update with new data
death.recent <- death.save[,c("Province/State","Country/Region","Lat","Long","3/14/20")]
data.recent <- data.save[,c("Province/State","Country/Region","Lat","Long","3/14/20")]


colnames(death.recent)[colnames(death.recent)=="3/14/20"] <- "count"
colnames(data.recent)[colnames(data.recent)=="3/14/20"] <- "count"




# World set-up credit to source here: 
#https://sarahleejane.github.io/learning/r/2014/09/21/plotting-data-points-on-maps-with-r.html

world.map <- map_data("world")

p <- ggplot() + coord_fixed() +
  xlab("") + ylab("")

#Add map to base plot
base.world <- p + geom_polygon(data=world.map, aes(x=long, y=lat, group=group), 
                               colour="#aeb3b0", fill="#aeb3b0")

theme <- 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_rect(fill = 'white', colour = 'white'), 
        axis.line = element_line(colour = "white"), legend.position="none",
        axis.ticks=element_blank(), axis.text.x=element_blank(),
        axis.text.y=element_blank())

world <- base.world + theme

COVID.world.case <- 
  world +
  geom_point(data=(data.recent %>% filter(count != 0)),
             aes(x=Long, y=Lat, size=count), colour="red", 
             fill="pink",pch=21, alpha=I(0.7)) +
             labs(title="COVID-19 Confirmed Cases", 
                  subtitle = "Point size according to number of cases")



COVID.world.death <- 
  world +
  geom_point(data=(death.recent %>% filter(count != 0)),
             aes(x=Long, y=Lat, size=count), colour="red", 
             fill="pink",pch=21, alpha=I(0.7)) +
             labs(title="COVID-19 Deaths", 
                  subtitle = "Point size according to number of deaths")

```

## World Map of COVID-19 {.tabset .tabset-fade}

Point size reflects number of cases / deaths

### Worldwide Confirmed Cases

```{r, echo=FALSE, fig.height=8, fig.width=12}

plot(COVID.world.case)

```

### Worldwide COVID-19 Deaths

```{r, echo=FALSE, fig.height=8, fig.width=12}

plot(COVID.world.death)

```


```{r, echo=FALSE, include=FALSE}


usa.map <- map_data("state")

#Add map to base plot (base plot created in previous section for world plot)
usa.map <- p + geom_polygon(data=usa.map, aes(x=long, y=lat, group=group), 
                               colour="#aeb3b0", fill="#aeb3b0")

#Apply the same mapping theme as the world map
usa <- usa.map + theme

# For the US cases I've specified the longitude and latitude in the filter to remove hawaii, alaska, puerto rico and limit to the continental US. 

usa.cases <- 
  usa +
  geom_point(data=(data.recent %>% 
                   filter(`Country/Region` %in% "US" & count != 0 & between(Long, -152,-50) & between(Lat, 27,49))), 
             aes(x=Long, y=Lat, size=count), colour="red", 
             fill="pink",pch=21, alpha=I(0.7)) +
             labs(title="COVID-19 Cases in the US", 
                  subtitle = "Point size according to number of cases")


usa.deaths <- 
  usa +
  geom_point(data=(death.recent %>% 
                   filter(`Country/Region` %in% "US" & count != 0 & between(Long, -152,-50) & between(Lat, 27,49))), 
             aes(x=Long, y=Lat, size=count), colour="red", 
             fill="pink",pch=21, alpha=I(0.7)) +
             labs(title="COVID-19 Deaths in the US", 
                  subtitle = "Point size according to number of deaths")


test <- data.recent %>% filter(`Country/Region` %in% "US" & count != 0 & between(Long, -152,-50) & between(Lat, 27,49)) 


```

## US Map of COVID-19 {.tabset .tabset-fade}

### US Confirmed Cases

```{r, echo=FALSE}

plot(usa.cases)

```

### US COVID-19 Deaths

```{r, echo=FALSE}

plot(usa.deaths)

```











