---
title: "COVID-19"
author: "Mike"
date: "3/17/2020"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Welcome to my first kaggle submission! Check back for weekly updates as this is very much a work in progress. Feedback, tips, and comments are appreciated. I hope you find this helpful!

```{r, echo=FALSE,include=FALSE}

library(tidyverse)
library(lubridate)
library(reshape2)
library(viridis)
library(ggthemes)
library(plotly)
library(maps)
library(formattable)
library(directlabels)

###################################
## Must update with data refresh ##
###################################

# This variable is referenced through the code, I found that the most recent data provided wasn't always complete
# therefore, setting this date will adjust the graphs and subsets throughout to use the most complete data availible
# Specified by you

latest_date <- list(as.Date('2020-03-14'))

# Setting these colors for use in the formattable table below
customGreen0 = "#DeF7E9"
customGreen = "#71CA97"
customRed = "#ff7f7f"
customYel = "#C9C271"


#########################
## Confirmed Cases Data #
#########################

#data <- read_csv("../input/novel-corona-virus-2019-dataset/time_series_covid_19_confirmed.csv")
data <- read_csv("time_series_covid_19_confirmed.csv")
data.save <- data
data <- data[,-c(1,3,4)]
data$`Country/Region` <- ifelse(grepl("Mainland China", data$`Country/Region`),'China',
                         ifelse(grepl("Viet Nam", data$`Country/Region`),'Vietnam',
                         ifelse(grepl("Taiwan*", data$`Country/Region`),'Taiwan', 
                         ifelse(grepl("Iran (Islamic Republic of)", data$`Country/Region`),'Iran',
                         ifelse(grepl("Russian Federation", data$`Country/Region`),'Russia',
                         ifelse(grepl("occupied Palestinian territory", data$`Country/Region`),'Palestine',
                         data$`Country/Region`))))))
# Summarise by Country
data <- data %>%
          group_by(`Country/Region`) %>% 
          summarise_all(list(sum))

# Melt into long form
data <- melt(data)

# Rename columns
colnames(data)[colnames(data)=="variable"] <- "date"
colnames(data)[colnames(data)=="value"] <- "count"

## Date Parsing shown below

###############
## Death Data #
###############

#death <- read_csv("../input/novel-corona-virus-2019-dataset/time_series_covid_19_deaths.csv")
death <- read_csv("time_series_covid_19_deaths.csv")
death.save <- death
death <- death[,-c(1,3,4)]
death$`Country/Region` <- ifelse(grepl("Mainland China", death$`Country/Region`),'China',
                         ifelse(grepl("Viet Nam", death$`Country/Region`),'Vietnam',
                         ifelse(grepl("Taiwan*", death$`Country/Region`),'Taiwan', 
                         ifelse(grepl("Iran (Islamic Republic of)", death$`Country/Region`),'Iran',
                         ifelse(grepl("Russian Federation", death$`Country/Region`),'Russia',
                         ifelse(grepl("occupied Palestinian territory", death$`Country/Region`),'Palestine',
                         death$`Country/Region`))))))

# Sum count by country
death <- death %>%
          group_by(`Country/Region`) %>% 
          summarise_all(list(sum))
# Melt into long form
death <- melt(death)
# Rename columns
colnames(death)[colnames(death)=="variable"] <- "date"
colnames(death)[colnames(death)=="value"] <- "count"
# Date to char for parsing and conversion
death$date <- as.character(death$date)
# Parse the date
death$year<-sapply(death$date, function(x) as.numeric(strsplit(x,"/")[[1]][3]))
death$month<-sapply(death$date, function(x) as.numeric(strsplit(x,"/")[[1]][1]))
death$day<-sapply(death$date, function(x) as.numeric(strsplit(x,"/")[[1]][2]))
# Put us back in the year 2000
death$year <- (death$year + 2000)
# Reformat date into y-m-d 
death$date<-as.Date(paste0(death$year,'-',death$month,'-',death$day), format="%Y-%m-%d")
# Add some weekdays for good measure
death$weekday <- weekdays(as.Date(death$date))
# Put weekdays in order
death$weekday <- ordered(death$weekday, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

##################
## Recovery Data #
##################

reco <- read_csv("time_series_covid_19_recovered.csv")
reco.save <- reco
reco <- reco[,-c(1,3,4)]
reco$`Country/Region` <- ifelse(grepl("Mainland China", reco$`Country/Region`),'China',
                         ifelse(grepl("Viet Nam", reco$`Country/Region`),'Vietnam',
                         ifelse(grepl("Taiwan*", reco$`Country/Region`),'Taiwan', 
                         ifelse(grepl("Iran (Islamic Republic of)", reco$`Country/Region`),'Iran',
                         ifelse(grepl("Russian Federation", reco$`Country/Region`),'Russia',
                         ifelse(grepl("occupied Palestinian territory", reco$`Country/Region`),'Palestine',
                         reco$`Country/Region`))))))

# Sum count by country
reco <- reco %>%
          group_by(`Country/Region`) %>% 
          summarise_all(list(sum))
# Melt into long form
reco <- melt(reco)
# Rename columns
colnames(reco)[colnames(reco)=="variable"] <- "date"
colnames(reco)[colnames(reco)=="value"] <- "count"
# Date to char for parsing and conversion
reco$date <- as.character(reco$date)
# Parse the date
reco$year<-sapply(reco$date, function(x) as.numeric(strsplit(x,"/")[[1]][3]))
reco$month<-sapply(reco$date, function(x) as.numeric(strsplit(x,"/")[[1]][1]))
reco$day<-sapply(reco$date, function(x) as.numeric(strsplit(x,"/")[[1]][2]))
# Put us back in the year 2000
reco$year <- (reco$year + 2000)
# Reformat date into y-m-d 
reco$date<-as.Date(paste0(reco$year,'-',reco$month,'-',reco$day), format="%Y-%m-%d")
# Add some weekdays for good measure
reco$weekday <- weekdays(as.Date(reco$date))
# Put weekdays in order
reco$weekday <- ordered(reco$weekday, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))


```

### Date Parsing and Formatting

Including the code below as it's nearly boilerplate for date formatting. In some cases the strsplit will need to be adjusted, but I find it much more efficient to parse and reformat a date each time than try to accommodate different date formats across datasets. 

```{r}

# Date to char for parsing and conversion
data$date <- as.character(data$date)

# Parse the date
data$year<-sapply(data$date, function(x) as.numeric(strsplit(x,"/")[[1]][3]))
data$month<-sapply(data$date, function(x) as.numeric(strsplit(x,"/")[[1]][1]))
data$day<-sapply(data$date, function(x) as.numeric(strsplit(x,"/")[[1]][2]))

# Put us back in the year 2000
data$year <- (data$year + 2000)

# Reformat date into y-m-d 
data$date<-as.Date(paste0(data$year,'-',data$month,'-',data$day), format="%Y-%m-%d")

# Add some weekdays for good measure
data$weekday <- weekdays(as.Date(data$date))

# Put weekdays in order
data$weekday <- ordered(data$weekday, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

# Set X axis limits 
time <- as.POSIXct(strptime(c("2020-01-22",latest_date), format = "%y-%m-%d"))

```

```{r, echo=FALSE, include=FALSE}
p <- ggplot(data, aes(date, `Country/Region`)) + 
     geom_tile(aes(fill = count)) +
     scale_fill_viridis(name="Confirmed Cases", option ="A", limits = c(0, 20000), oob = scales::squish) +
     labs(title = "Global Confirmed COVID-19 Cases by Country", subtitle = "Count of cases")

```

## Time Series Heatmap 

At present, with China as a large confirmed case outlier (over 80,000), I rescaled the confirmed cases using a limit of 10,000 to give greater visibility to lower and emerging counts. 

```{r, echo=FALSE, fig.height=20, fig.width=18}

plot(p)


```


```{r, echo=FALSE, include=FALSE}

# Creating a new subset of the top 10 countries by cases
top.ten <- data %>% filter(date == latest_date) %>% arrange(desc(count)) %>% top_n(10, count)
# subsetting the death and reco dataframes by latest date
tbl.all <- top.ten
tbl.death <- death %>% filter(date == latest_date)
tbl.reco <- reco %>% filter(date == latest_date)

# Joining the top 10 countries by cases to their respective counts of deaths and recoveries
tbl.all <- tbl.all %>% left_join(tbl.death[,c("count", "Country/Region")], by=c("Country/Region"), suffix = c("","_death"))
tbl.all <- tbl.all %>% left_join(tbl.reco[,c("count", "Country/Region")], by=c("Country/Region"), suffix = c("","_reco"))

#renaming the columns
colnames(tbl.all)[colnames(tbl.all)=="count"] <- "Cases"
colnames(tbl.all)[colnames(tbl.all)=="count_death"] <- "Deaths"
colnames(tbl.all)[colnames(tbl.all)=="count_reco"] <- "Recoveries"

tbl.all <- subset(tbl.all, select = c(`Country/Region`, Cases, Deaths, Recoveries))

tbl.all$`Percent Recover` <- NA
tbl.all$Mortality <- NA

# Calculating Percent
tbl.all$`Percent Recover` <- percent((tbl.all$Recoveries / tbl.all$Cases))
tbl.all$Mortality <- percent((tbl.all$Deaths / tbl.all$Cases)) 

# Formattable is pretty neat for making nicer charts
table.counts <- formattable(tbl.all, align =c("l","c","c","c","c","c"),
                list(`SDOH` = formatter("span", style = ~ formattable::style(color = "grey",font.weight = "bold")),
                `Cases` = color_bar(customYel),
                `Recoveries` = color_bar(customGreen),
                `Deaths` = color_bar(customRed),
                `Percent Recover` = color_bar(customGreen),
                `Mortality` = color_bar(customRed)))

```

<br>
<br>

## Cases, Recoveries, and Deaths

For the top 10 countries with the most cases, most recent recovery and death counts are included below. Recovery and Mortality are calculated from the datasets provided and therefore DO NOT represent an accurate representation of those rates which are still being determined. Additionally, we would expect a significant delay in reporting of recovered cases. 

```{r, echo=FALSE}

formattable(table.counts)

```


```{r, echo=FALSE, include=FALSE}

# Because the most recent date has missing data, I'm using the next most recent date
top <- data %>% filter(date == latest_date) %>% arrange(desc(count)) %>% top_n(20, count)


trend <- data %>% filter(`Country/Region` %in% top$`Country/Region`) %>%
                  group_by(date, count) %>%
                  ggplot(aes(x=as.POSIXct(date),y=count, color=`Country/Region`)) + 
                  geom_point(size=.5) + 
                  geom_line(alpha=.5) + 
                  scale_x_datetime(limits =time) + 
                  scale_color_viridis(discrete = TRUE, option ="A") +
                  labs(title="Total COVID-19 Confirmed Case Count") +
                  xlab("Date") + 
                  ylab("Cases")


lol.trend <- ggplot(top, aes(x=reorder(`Country/Region`,count), y=count)) +
                  geom_segment( aes(x=reorder(`Country/Region`,count), xend=reorder(`Country/Region`,count), y=0, yend=count)) +
                  geom_point(size=5, color="red", fill=alpha("pink", 0.3), alpha=0.7, shape=21, stroke=2) +
                  coord_flip() +
                  labs(x="Country", y="Count") +
                  labs(title="Top 20 Countries Confirmed Case Count")

```

<br>
<br>

## Confirmed Cases Trend {.tabset .tabset-fade}

### Top 20 Global Trend

Using ggplotly for an interactive plot, here are the Country/Region's in the top 20 for confirmed cases plotted since the pandemic started.

```{r, echo=FALSE, fig.height=8, fig.width=10}

ggplotly(trend)

```

### Top 20 Global Count

```{r, echo=FALSE, fig.height=8, fig.width=10}

plot(lol.trend)

```


```{r, echo=FALSE, include=FALSE}


top.death <- death %>% filter(date == latest_date) %>% arrange(desc(count)) %>% top_n(20, count)

trend.death <- death %>% filter(`Country/Region` %in% top.death$`Country/Region`) %>%
                  group_by(date, count) %>%
                  ggplot(aes(x=as.POSIXct(date),y=count, color=`Country/Region`)) + 
                  geom_point(size=.5) + 
                  geom_line(alpha=.5) + 
                  scale_x_datetime(limits =time) + 
                  scale_color_viridis(discrete = TRUE, option ="A") +
                  labs(title="Global COVID-19 Deaths") +
                  xlab("Date") + 
                  ylab("Deaths")


lol.death <- ggplot(top.death, aes(x=reorder(`Country/Region`,count), y=count)) +
                  geom_segment( aes(x=reorder(`Country/Region`,count), xend=reorder(`Country/Region`,count), y=0, yend=count)) +
                  geom_point(size=5, color="red", fill=alpha("pink", 0.3), alpha=0.7, shape=21, stroke=2) +
                  coord_flip() +
                  labs(x="Country", y="Count") +
                  labs(title="Top 20 Countries Death Count")

```

<br>
<br>

## Death Trend {.tabset .tabset-fade}

### Top 20 Global Death Trend

Time series plot of top 20 County/Region's with COVID-19 related deaths. There may be more than 20 countries listed depending on how many are "tied" for the 20th spot (to be fair, its an arbitrary cutoff for the visualization).

```{r, echo=FALSE, fig.height=8, fig.width=10}

ggplotly(trend.death)

```

### Top 20 Global Death Count


```{r, echo=FALSE, fig.height=8, fig.width=10}

plot(lol.death)

```

```{r, echo=FALSE, include=FALSE}

# Subset of countries with 100 cases arranged in ascending date order, Then an ID column is created to increment 1 day by country to get the number of days since 100 confirmed cases between(Long, -152,-50)
hun.cases <- data %>% filter(count >= 100) %>% arrange(date)
#hun.cases <- data %>% filter(between(count, 100,40000)) %>% arrange(date)
hun.cases <-hun.cases %>% group_by(`Country/Region`) %>% mutate(id = row_number())
hun.cases <-hun.cases %>% group_by(`Country/Region`) %>% mutate(id = 1:n())
hun.cases <-hun.cases %>% group_by(`Country/Region`) %>% mutate(id = seq_len(n()))
hun.cases <-hun.cases %>% group_by(`Country/Region`) %>% mutate(id = seq_along(date))




# plotting the top 10 countries by number of cases
# direct labels is used here to add the country name to the end of the line
hun.trend <- hun.cases %>% filter(`Country/Region` %in% top.ten$`Country/Region`) %>%
                  group_by(id, count) %>%
                  ggplot(aes(x=id,y=count, color=`Country/Region`)) + 
                  geom_point(size=.5) + 
                  geom_line(alpha=.9) +
                  ylim(0,40000) + 
                  xlim(0,30) +
                  scale_color_viridis(discrete = TRUE, option ="A") +
                  labs(title="Case Trend Since 100th Case",
                       subtitle = "Top 10 countries by case count") +
                  xlab("Number of Days") + 
                  ylab("Case Count")


```

<br>
<br>

## Trend since first 100

Plotting the trend of the top 10 countries by case, number of days since their initial 100 confirmed cases. This gives us a sense of the effectiveness of a countries response in containing the virus. Again, since with China in more advanced stages of the pandemic the x and y axis have been limited.  

```{r, echo=FALSE, warning=FALSE, fig.height=8, fig.width=10}

ggplotly(hun.trend)

```

<br>
<br>

## Cases, Recoveries and Deaths (by country) {.tabset .tabset-fade}

Graphs of relationship between confirmed cases and death for highly impacted countries. 

### China

Example code included, overlaying two area plots from seperate data frames. I wasn't able to identify a straightforward method to include a plot legend-- comment below if you have one!

```{r}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "China") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customYel, alpha=0.4, fill = customYel) +
      geom_area(data=(reco %>% 
                  filter(`Country/Region` %in% "China") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customGreen, alpha=0.4, fill = customGreen) +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "China") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customRed, fill=customRed) +
                  labs(title="China Confirmed Cases and Deaths", 
                       subtitle = " Yellow is confirmed cases \n Green is number of recoveries \n Red is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```


### US

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "US") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customYel, alpha=0.4, fill = customYel) +
      geom_area(data=(reco %>% 
                  filter(`Country/Region` %in% "US") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customGreen, alpha=0.4, fill = customGreen) +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "US") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customRed, fill=customRed) +
                  labs(title="US Confirmed Cases and Deaths", 
                       subtitle = " Yellow is confirmed cases \n Green is number of recoveries \n Red is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```

### Italy

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "Italy") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customYel, alpha=0.4, fill = customYel) +
      geom_area(data=(reco %>% 
                  filter(`Country/Region` %in% "Italy") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customGreen, alpha=0.4, fill = customGreen) +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "Italy") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customRed, fill=customRed) +
                  labs(title="Italy Confirmed Cases and Deaths", 
                       subtitle = " Yellow is confirmed cases \n Green is number of recoveries \n Red is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```


### Iran

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "Iran") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customYel, alpha=0.4, fill = customYel) +
      geom_area(data=(reco %>% 
                  filter(`Country/Region` %in% "Iran") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customGreen, alpha=0.4, fill = customGreen) +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "Iran") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customRed, fill=customRed) +
                  labs(title="Iran Confirmed Cases and Deaths", 
                       subtitle = " Yellow is confirmed cases \n Green is number of recoveries \n Red is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```

### South Korea

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "Korea, South") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customYel, alpha=0.4, fill = customYel) +
      geom_area(data=(reco %>% 
                  filter(`Country/Region` %in% "Korea, South") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customGreen, alpha=0.4, fill = customGreen) +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "Korea, South") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customRed, fill=customRed) +
                  labs(title="South Korea Confirmed Cases and Deaths", 
                       subtitle = " Yellow is confirmed cases \n Green is number of recoveries \n Red is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```


### Spain

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "Spain") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customYel, alpha=0.4, fill = customYel) +
      geom_area(data=(reco %>% 
                  filter(`Country/Region` %in% "Spain") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customGreen, alpha=0.4, fill = customGreen) +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "Spain") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customRed, fill=customRed) +
                  labs(title="Spain Confirmed Cases and Deaths", 
                       subtitle = " Yellow is confirmed cases \n Green is number of recoveries \n Red is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```


```{r, echo=FALSE, include=FALSE}

#-----------------------------------------------------------------------------------------------Must update with new data
death.recent <- death.save[,c("Province/State","Country/Region","Lat","Long","3/14/20")]
data.recent <- data.save[,c("Province/State","Country/Region","Lat","Long","3/14/20")]


colnames(death.recent)[colnames(death.recent)=="3/14/20"] <- "count"
colnames(data.recent)[colnames(data.recent)=="3/14/20"] <- "count"




# World set-up credit to source here: 
#https://sarahleejane.github.io/learning/r/2014/09/21/plotting-data-points-on-maps-with-r.html

world.map <- map_data("world")

p <- ggplot() + coord_fixed() +
  xlab("") + ylab("")

#Add map to base plot
base.world <- p + geom_polygon(data=world.map, aes(x=long, y=lat, group=group), 
                               colour="#aeb3b0", fill="#aeb3b0")

theme <- 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_rect(fill = 'white', colour = 'white'), 
        axis.line = element_line(colour = "white"), legend.position="none",
        axis.ticks=element_blank(), axis.text.x=element_blank(),
        axis.text.y=element_blank())

world <- base.world + theme

COVID.world.case <- 
  world +
  geom_point(data=(data.recent %>% filter(count != 0)),
             aes(x=Long, y=Lat, size=count), colour="red", 
             fill="pink",pch=21, alpha=I(0.7)) +
             labs(title="COVID-19 Confirmed Cases", 
                  subtitle = "Point size according to number of cases")



COVID.world.death <- 
  world +
  geom_point(data=(death.recent %>% filter(count != 0)),
             aes(x=Long, y=Lat, size=count), colour="red", 
             fill="pink",pch=21, alpha=I(0.7)) +
             labs(title="COVID-19 Deaths", 
                  subtitle = "Point size according to number of deaths")

```

## World Map {.tabset .tabset-fade}

Point size reflects number of cases / deaths

### Worldwide Confirmed Cases

```{r, echo=FALSE, fig.height=8, fig.width=12}

plot(COVID.world.case)

```

### Worldwide COVID-19 Deaths

```{r, echo=FALSE, fig.height=8, fig.width=12}

plot(COVID.world.death)

```


```{r, echo=FALSE, include=FALSE}


usa.map <- map_data("state")

#Add map to base plot (base plot created in previous section for world plot)
usa.map <- p + geom_polygon(data=usa.map, aes(x=long, y=lat, group=group), 
                               colour="#aeb3b0", fill="#aeb3b0")

#Apply the same mapping theme as the world map
usa <- usa.map + theme

# For the US cases I've specified the longitude and latitude in the filter to remove hawaii, alaska, puerto rico and limit to the continental US. 

usa.cases <- 
  usa +
  geom_point(data=(data.recent %>% 
                   filter(`Country/Region` %in% "US" & count != 0 & between(Long, -152,-50) & between(Lat, 27,49))), 
             aes(x=Long, y=Lat, size=count), colour="red", 
             fill="pink",pch=21, alpha=I(0.7)) +
             labs(title="COVID-19 Cases in the US", 
                  subtitle = "Point size according to number of cases")


usa.deaths <- 
  usa +
  geom_point(data=(death.recent %>% 
                   filter(`Country/Region` %in% "US" & count != 0 & between(Long, -152,-50) & between(Lat, 27,49))), 
             aes(x=Long, y=Lat, size=count), colour="red", 
             fill="pink",pch=21, alpha=I(0.7)) +
             labs(title="COVID-19 Deaths in the US", 
                  subtitle = "Point size according to number of deaths")




```

## US Map {.tabset .tabset-fade}

### US Confirmed Cases

```{r, echo=FALSE}

plot(usa.cases)

```

### US COVID-19 Deaths

```{r, echo=FALSE}

plot(usa.deaths)

```

