---
title: "COVID-19 Tracking"
author: "Mike"
date: "3/17/2020"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Welcome to my first kaggle submission! Check back for weekly updates as this is very much a work in progress. Feedback, tips, and comments are appreciated. I hope you find this helpful!

Update 3/23/20: Updated Maps using Leaflet, US Choropleth and per capita case and death rates.


```{r, echo=FALSE,include=FALSE}

library(tidyverse)
library(lubridate)
library(reshape2)
library(viridis)
library(ggthemes)
library(plotly)
library(maps)
library(formattable)
library(directlabels)
library(leaflet)
library(scales)
library(usmap)

###################################
## Must update with data refresh ##
###################################

# This variable is referenced through the code, I found that the most recent data provided wasn't always complete
# therefore, setting this date will adjust the graphs and subsets throughout to use the most complete data availible
# Specified by you

latest_date <- list(as.Date('2020-03-22'))

# Setting these colors for use in the formattable table below
customGreen0 = "#DeF7E9"
customGreen = "#71CA97"
customRed = "#ff7f7f"
customYel = "#C9C271"

#########################
## Data New             #
#########################

data <- read_csv("covid_19_data.csv")
#data <- read_csv("../input/novel-corona-virus-2019-dataset/covid_19_data.csv")

data.save <- data

data <- data[,-c(1,3,5)]
data$`Country/Region` <- ifelse(grepl("Mainland China", data$`Country/Region`),'China',
                         ifelse(grepl("occupied Palestinian territory", data$`Country/Region`),'Palestine',
                         data$`Country/Region`))


###############
## Death Data #
###############

death <- subset(data, select = c(`Country/Region`, ObservationDate, Deaths))
colnames(death)[colnames(death)=="ObservationDate"] <- "date"
colnames(death)[colnames(death)=="Deaths"] <- "count"

death <- death %>%
          group_by(`Country/Region`,date) %>% 
          summarise_all(list(sum))


# Date to char for parsing and conversion
death$date <- as.character(death$date)
# Parse the date
death$year<-sapply(death$date, function(x) as.numeric(strsplit(x,"/")[[1]][3]))
death$month<-sapply(death$date, function(x) as.numeric(strsplit(x,"/")[[1]][1]))
death$day<-sapply(death$date, function(x) as.numeric(strsplit(x,"/")[[1]][2]))

# Some years are "2020" some are "20" so this will set them all straight :)
death$year <- ifelse(death$year == '2020', "20", death$year)
# Put us back in the year 2000
death$year <- (as.numeric(death$year) + 2000)
# Reformat date into y-m-d 
death$date<-as.Date(paste0(death$year,'-',death$month,'-',death$day), format="%Y-%m-%d")
# Add some weekdays for good measure
death$weekday <- weekdays(as.Date(death$date))
# Put weekdays in order
death$weekday <- ordered(death$weekday, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))


##################
## Recovery Data #
##################

reco <- subset(data, select = c(`Country/Region`, ObservationDate, Recovered))
colnames(reco)[colnames(reco)=="ObservationDate"] <- "date"
colnames(reco)[colnames(reco)=="Recovered"] <- "count"

reco <- reco %>%
          group_by(`Country/Region`,date) %>% 
          summarise_all(list(sum))

# Date to char for parsing and conversion
reco$date <- as.character(reco$date)
# Parse the date
reco$year<-sapply(reco$date, function(x) as.numeric(strsplit(x,"/")[[1]][3]))
reco$month<-sapply(reco$date, function(x) as.numeric(strsplit(x,"/")[[1]][1]))
reco$day<-sapply(reco$date, function(x) as.numeric(strsplit(x,"/")[[1]][2]))

# Some years are "2020" some are "20" so this will set them all straight :)
reco$year <- ifelse(reco$year == '2020', "20", reco$year)
# Put us back in the year 2000
reco$year <- (as.numeric(reco$year) + 2000)
# Reformat date into y-m-d 
reco$date<-as.Date(paste0(reco$year,'-',reco$month,'-',reco$day), format="%Y-%m-%d")
# Add some weekdays for good measure
reco$weekday <- weekdays(as.Date(reco$date))
# Put weekdays in order
reco$weekday <- ordered(reco$weekday, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

#########################
## Confirmed Cases Data #
#########################

data <- subset(data, select = c(`Country/Region`, ObservationDate, Confirmed))
colnames(data)[colnames(data)=="ObservationDate"] <- "date"
colnames(data)[colnames(data)=="Confirmed"] <- "count"

data <- data %>%
          group_by(`Country/Region`,date) %>% 
          summarise_all(list(sum))

#Date Parsing Happens below:


```

## Date Parsing and Formatting {.tabset .tabset-fade}

### Note

Including the date formatting code on the next tab as it's nearly boilerplate. In some cases the "strsplit" will need to be adjusted but I find it much more efficient to parse and reformat the date each time than try to accommodate different date formats across different datasets. 

### Code

```{r}

# Date to char for parsing and conversion
data$date <- as.character(data$date)
# Parse the date
data$year<-sapply(data$date, function(x) as.numeric(strsplit(x,"/")[[1]][3]))
data$month<-sapply(data$date, function(x) as.numeric(strsplit(x,"/")[[1]][1]))
data$day<-sapply(data$date, function(x) as.numeric(strsplit(x,"/")[[1]][2]))

# Some years are "2020" some are "20" so this will set them all straight :)
data$year <- ifelse(data$year == '2020', "20", data$year)
# Put us back in the year 2000
data$year <- (as.numeric(data$year) + 2000)
# Reformat date into y-m-d 
data$date<-as.Date(paste0(data$year,'-',data$month,'-',data$day), format="%Y-%m-%d")
# Add some weekdays for good measure
data$weekday <- weekdays(as.Date(data$date))
# Put weekdays in order
data$weekday <- ordered(data$weekday, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

# Set X axis limits 
time <- as.POSIXct(strptime(c("2020-01-22",latest_date), format = "%y-%m-%d"))

```

```{r, echo=FALSE, include=FALSE}
p <- ggplot(data, aes(date, `Country/Region`)) + 
     geom_tile(aes(fill = count)) +
     scale_fill_viridis(name="Confirmed Cases", 
                        option ="A", 
                        limits = c(0, 20000), 
                        oob = scales::squish) +
     theme(panel.background=element_rect(fill="black", color="black"),
           plot.background = element_blank(),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           panel.border = element_blank()) +
     labs(title = "Global Confirmed COVID-19 Cases by Country", 
          subtitle = "Count of cases")



```

## Time Series Heatmap 

At present, with China as a large confirmed case outlier (over 80,000), I rescaled the confirmed cases using a limit of 10,000 to give greater visibility to lower and emerging counts. 

```{r, echo=FALSE, fig.height=30, fig.width=18}

plot(p)


```


```{r, echo=FALSE, include=FALSE, warning=FALSE}

# Creating a new subset of the top 10 countries by cases

top.ten <- data %>% filter(date == latest_date) %>% arrange(desc(count)) 
top.ten <- top_n(ungroup(top.ten), 10, count)

# subsetting the death and reco dataframes by latest date
tbl.all <- top.ten
tbl.death <- death %>% filter(date == latest_date)
tbl.reco <- reco %>% filter(date == latest_date)

# Joining the top 10 countries by cases to their respective counts of deaths and recoveries
tbl.all <- tbl.all %>% left_join(tbl.death[,c("count", "Country/Region")], by=c("Country/Region"), suffix = c("","_death"))
tbl.all <- tbl.all %>% left_join(tbl.reco[,c("count", "Country/Region")], by=c("Country/Region"), suffix = c("","_reco"))

#renaming the columns
colnames(tbl.all)[colnames(tbl.all)=="count"] <- "Cases"
colnames(tbl.all)[colnames(tbl.all)=="count_death"] <- "Deaths"
colnames(tbl.all)[colnames(tbl.all)=="count_reco"] <- "Recoveries"

tbl.all <- subset(tbl.all, select = c(`Country/Region`, Cases, Deaths, Recoveries))

tbl.all$`Percent Recover` <- NA
tbl.all$Mortality <- NA

# Calculating Percent
tbl.all$`Percent Recover` <- percent((tbl.all$Recoveries / tbl.all$Cases))
tbl.all$Mortality <- percent((tbl.all$Deaths / tbl.all$Cases)) 

# Formattable is pretty neat for making nicer charts
table.counts <- formattable(tbl.all, align =c("l","r","r","r","r","r"),
                list(`SDOH` = formatter("span", style = ~ formattable::style(color = "grey",font.weight = "bold")),
                `Cases` = color_bar(customYel),
                `Recoveries` = color_bar(customGreen),
                `Deaths` = color_bar(customRed),
                `Percent Recover` = color_bar(customGreen),
                `Mortality` = color_bar(customRed)))

```

<br>
<br>

## Cases, Recoveries, and Deaths

For the top 10 countries with the most cases, most recent recovery and death counts are included below. Recovery and Mortality are calculated from the datasets provided. Additionally, we would expect a significant delay in reporting of recovered cases. 

```{r, echo=FALSE, warning=FALSE}

formattable(table.counts)

```


```{r, echo=FALSE, include=FALSE}

# Because the most recent date has missing data, I'm using the next most recent date
top <- data %>% filter(date == latest_date) %>% arrange(desc(count))
top <- top_n(ungroup(top), 20, count)

trend <- data %>% filter(`Country/Region` %in% top$`Country/Region`) %>%
                  group_by(date, count) %>%
                  ggplot(aes(x=as.POSIXct(date),y=count, color=`Country/Region`)) + 
                  geom_point(size=.5) + 
                  geom_line(alpha=.5) + 
                  scale_x_datetime(limits =time) + 
                  scale_color_viridis(discrete = TRUE, option ="A") +
                  labs(title="Total COVID-19 Confirmed Case Count") +
                  xlab("Date") + 
                  ylab("Cases")


lol.trend <- ggplot(top, aes(x=reorder(`Country/Region`,count), y=count)) +
                  geom_segment( aes(x=reorder(`Country/Region`,count), xend=reorder(`Country/Region`,count), y=0, yend=count)) +
                  geom_point(size=5, color="red", fill=alpha("pink", 0.3), alpha=0.7, shape=21, stroke=2) +
                  coord_flip() +
                  labs(x="Country", y="Count") +
                  labs(title="Top 20 Countries Confirmed Case Count")

```

<br>
<br>

## Confirmed Cases Trend {.tabset .tabset-fade}

### Top 20 Global Trend

Using ggplotly for an interactive plot, here are the Country/Region's in the top 20 for confirmed cases plotted since the pandemic started.

```{r, echo=FALSE, fig.height=8, fig.width=10}

ggplotly(trend)

```

### Top 20 Global Count

```{r, echo=FALSE, fig.height=8, fig.width=10}

plot(lol.trend)

```


```{r, echo=FALSE, include=FALSE}


top.death <- death %>% filter(date == latest_date) %>% arrange(desc(count))
top.death <- top_n(ungroup(top.death), 20, count)


trend.death <- death %>% filter(`Country/Region` %in% top.death$`Country/Region`) %>%
                  group_by(date, count) %>%
                  ggplot(aes(x=as.POSIXct(date),y=count, color=`Country/Region`)) + 
                  geom_point(size=.5) + 
                  geom_line(alpha=.5) + 
                  scale_x_datetime(limits =time) + 
                  scale_color_viridis(discrete = TRUE, option ="A") +
                  labs(title="Global COVID-19 Deaths") +
                  xlab("Date") + 
                  ylab("Deaths")


lol.death <- ggplot(top.death, aes(x=reorder(`Country/Region`,count), y=count)) +
                  geom_segment( aes(x=reorder(`Country/Region`,count), xend=reorder(`Country/Region`,count), y=0, yend=count)) +
                  geom_point(size=5, color="red", fill=alpha("pink", 0.3), alpha=0.7, shape=21, stroke=2) +
                  coord_flip() +
                  labs(x="Country", y="Count") +
                  labs(title="Top 20 Countries Death Count")

```

<br>
<br>

## Death Trend {.tabset .tabset-fade}

### Top 20 Global Death Trend

Time series plot of top 20 County/Region's with COVID-19 related deaths. There may be more than 20 countries listed depending on how many are "tied" for the 20th spot (to be fair, its an arbitrary cutoff for the visualization).

```{r, echo=FALSE, fig.height=8, fig.width=10}

ggplotly(trend.death)

```

### Top 20 Global Death Count


```{r, echo=FALSE, fig.height=8, fig.width=10}

plot(lol.death)

```

```{r, echo=FALSE, include=FALSE}

# Subset of countries with 100 cases arranged in ascending date order, Then an ID column is created to increment 1 day by country to get the number of days since 100 confirmed cases between(Long, -152,-50)
hun.cases <- data %>% filter(count >= 100) %>% arrange(date)
#hun.cases <- data %>% filter(between(count, 100,40000)) %>% arrange(date)
hun.cases <-hun.cases %>% group_by(`Country/Region`) %>% mutate(id = row_number())
hun.cases <-hun.cases %>% group_by(`Country/Region`) %>% mutate(id = 1:n())
hun.cases <-hun.cases %>% group_by(`Country/Region`) %>% mutate(id = seq_len(n()))
hun.cases <-hun.cases %>% group_by(`Country/Region`) %>% mutate(id = seq_along(date))




# plotting the top 10 countries by number of cases
# direct labels is used here to add the country name to the end of the line
hun.trend <- hun.cases %>% filter(`Country/Region` %in% top.ten$`Country/Region`) %>%
                  group_by(id, count) %>%
                  ggplot(aes(x=id,y=count, color=`Country/Region`)) + 
                  geom_point(size=.5) + 
                  geom_line(alpha=.9) +
                  ylim(0,40000) + 
                  xlim(0,30) +
                  scale_color_viridis(discrete = TRUE, option ="A") +
                  labs(title="Case Trend Since 100th Case",
                       subtitle = "Top 10 countries by case count") +
                  xlab("Number of Days") + 
                  ylab("Case Count")


```

<br>
<br>

## Trend since first 100

Plotting the trend of the top 10 countries by case, number of days since their initial 100 confirmed cases. This gives us a sense of the effectiveness of a countries response in containing the virus. Again, since with China in more advanced stages of the pandemic the x and y axis have been limited. Here we see US cases climing faster than some of the hardest hit countries China, Italy, and Iran.  

```{r, echo=FALSE, warning=FALSE, fig.height=8, fig.width=10}

ggplotly(hun.trend)

```

<br>
<br>

## Cases, Recoveries and Deaths (by country) {.tabset .tabset-fade}

Graphs of relationship between confirmed cases and death for highly impacted countries. 

### China

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "China") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customYel, alpha=0.4, fill = customYel) +
      geom_area(data=(reco %>% 
                  filter(`Country/Region` %in% "China") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customGreen, alpha=0.4, fill = customGreen) +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "China") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customRed, fill=customRed) +
                  labs(title="China Confirmed Cases and Deaths", 
                       subtitle = " Yellow is confirmed cases \n Green is number of recoveries \n Red is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```


### US

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "US") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customYel, alpha=0.4, fill = customYel) +
      geom_area(data=(reco %>% 
                  filter(`Country/Region` %in% "US") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customGreen, alpha=0.4, fill = customGreen) +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "US") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customRed, fill=customRed) +
                  labs(title="US Confirmed Cases and Deaths", 
                       subtitle = " Yellow is confirmed cases \n Green is number of recoveries \n Red is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```

### Italy

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "Italy") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customYel, alpha=0.4, fill = customYel) +
      geom_area(data=(reco %>% 
                  filter(`Country/Region` %in% "Italy") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customGreen, alpha=0.4, fill = customGreen) +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "Italy") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customRed, fill=customRed) +
                  labs(title="Italy Confirmed Cases and Deaths", 
                       subtitle = " Yellow is confirmed cases \n Green is number of recoveries \n Red is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```


### Iran

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "Iran") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customYel, alpha=0.4, fill = customYel) +
      geom_area(data=(reco %>% 
                  filter(`Country/Region` %in% "Iran") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customGreen, alpha=0.4, fill = customGreen) +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "Iran") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customRed, fill=customRed) +
                  labs(title="Iran Confirmed Cases and Deaths", 
                       subtitle = " Yellow is confirmed cases \n Green is number of recoveries \n Red is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```

### South Korea

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "South Korea") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customYel, alpha=0.4, fill = customYel) +
      geom_area(data=(reco %>% 
                  filter(`Country/Region` %in% "South Korea") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customGreen, alpha=0.4, fill = customGreen) +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "South Korea") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customRed, fill=customRed) +
                  labs(title="South Korea Confirmed Cases and Deaths", 
                       subtitle = " Yellow is confirmed cases \n Green is number of recoveries \n Red is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```


### Spain

```{r, echo=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "Spain") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customYel, alpha=0.4, fill = customYel) +
      geom_area(data=(reco %>% 
                  filter(`Country/Region` %in% "Spain") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customGreen, alpha=0.4, fill = customGreen) +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "Spain") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customRed, fill=customRed) +
                  labs(title="Spain Confirmed Cases and Deaths", 
                       subtitle = " Yellow is confirmed cases \n Green is number of recoveries \n Red is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```

### Code

Example code included, overlaying area plots from separate data frames. I wasn't able to identify a straightforward method to include a plot legend when combining multiple gg objects as done here-- comment below if you have one!

```{r, eval=FALSE}

ggplot() + 
      geom_area(data=(data %>% 
                  filter(`Country/Region` %in% "China") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customYel, alpha=0.4, fill = customYel) +
      geom_area(data=(reco %>% 
                  filter(`Country/Region` %in% "China") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customGreen, alpha=0.4, fill = customGreen) +
      geom_area(data=(death %>% 
                  filter(`Country/Region` %in% "China") %>%
                  group_by(date, count)), 
                  aes(x=as.POSIXct(date),y=count), color=customRed, fill=customRed) +
                  labs(title="Spain Confirmed Cases and Deaths", 
                       subtitle = " Yellow is confirmed cases \n Green is number of recoveries \n Red is number of deaths") +
                  xlab("Date") + 
                  ylab("Count of Cases / Deaths") 


```



```{r, echo=FALSE}


#---------------------------------------------------------------------------------------------DATA LOAD STEP


death.save <- read_csv("time_series_covid_19_deaths.csv")
data.save <- read_csv("time_series_covid_19_confirmed.csv")

#data.save <- read_csv("../input/novel-corona-virus-2019-dataset/time_series_covid_19_confirmed.csv")
#death.save <- read_csv("../input/novel-corona-virus-2019-dataset/time_series_covid_19_deaths.csv")

#---------------------------------------------------------------------------------------------DATE UPDATE

death.recent <- death.save[,c("Province/State","Country/Region","Lat","Long","3/22/20")]
data.recent <- data.save[,c("Province/State","Country/Region","Lat","Long","3/22/20")]


colnames(death.recent)[colnames(death.recent)=="3/22/20"] <- "count"
colnames(data.recent)[colnames(data.recent)=="3/22/20"] <- "count"

death.recent <- death.recent %>% filter(count > 0)
data.recent <- data.recent %>% filter(count > 0)

cases.death.world <- leaflet() %>% 
   setView( lng = -10, lat = 30, zoom = 2 ) %>% 
   addProviderTiles("Esri.WorldImagery", group="Satellite") %>%
   addTiles(options = providerTileOptions(noWrap = TRUE), group="Map") %>%

   addCircleMarkers(data=death.recent, 
                    lng=~Long,
                    lat=~Lat,
                    radius=(rescale(death.recent$count, to=c(0,10))), 
                    color="black",
                    fillColor="red", 
                    stroke = TRUE, 
                    popup = ~as.character(count), 
                    label = ~as.character(count), 
                    fillOpacity = 0.8, 
                    group="Deaths") %>%
  
   addCircleMarkers(data=data.recent, 
                    lng=~Long, 
                    lat=~Lat, 
                    radius=(rescale(data.recent$count, to=c(0,10))), 
                    color="black",
                    fillColor="yellow",
                    stroke = TRUE,
                    popup = ~as.character(count), 
                    label = ~as.character(count), 
                    fillOpacity = 0.8,
                    group="Cases") %>%
  
   addLayersControl(overlayGroups = c("Deaths","Cases"),
                    baseGroups = c("Map","Satellite"), 
                    options = layersControlOptions(collapsed = FALSE))





```

## Worldwide COVID Cases/Deaths

Using the leaflet package here to generate an interactive worldwide map of COVID 19 Confirmed cases and deaths. Markers are sized according to count of either cases or deaths scaled to between 0 and 10. 


```{r, echo=FALSE}

cases.death.world

```



```{r, echo=FALSE, include=FALSE}

#Loading package here as it will mask the "plot" object from base graphics.
library(tigris)

# Create the map object using the tigris package
states <- states(cb=T)

# filter down the current data for US states
count.state <- data.recent %>%
  filter(`Country/Region` == "US") %>%
  group_by(`Province/State`) %>%
  rename(state=`Province/State`,
         total=count)

# Merge the state counts into the geo data
state.merge <- geo_join(states, count.state, "NAME", "state")

# Adjust Rate per 100,000
# To adjust confirmed cases  relative to population: 
# load us map for 2015 Census population data
#Take the statepop from the us mappackage: 
state.pop <- statepop
colnames(state.pop)[colnames(state.pop)=="full"] <- "state"

# left join the data
count.state <- count.state %>% left_join(state.pop, by=c("state"))

# Calculating infections per 100,000 residents rounded to 2 digits
count.state$Confirmed_Per_Capita <- round(count.state$total/count.state$pop_2015*100000,2)

#merge updated counts into geo data
state.merge <- geo_join(states, count.state, "NAME", "state")

#Set the color palette (viridis package option a = Magma)
pal <- colorNumeric("magma", domain=state.merge$Confirmed_Per_Capita, na.color="transparent")

# Set the popup text option
popup_t <- paste0("State: ", as.character(state.merge$state),  "<br>", 
                  "Confirmed Cases Per 100,000 Residents: ", as.character(state.merge$Confirmed_Per_Capita),  "<br>",
                  "Total Cases: ", as.character(state.merge$total))


```

<br>
<br>

## US Choropleth Map

An interactive Choropleth map showing confirmed COVID-19 Cases per 100,000 residents (2015 US Census Data):

```{r, echo=FALSE}

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-95, 38, zoom = 4) %>% 
  addPolygons(data = state.merge , 
              fillColor = ~pal(state.merge$Confirmed_Per_Capita), 
              fillOpacity = 0.7, 
              weight = 0.2, 
              smoothFactor = 0.2, 
              popup = ~popup_t) %>%
  addLegend(pal = pal, 
            values = state.merge$Confirmed_Per_Capita, 
            position = "bottomright", 
            title = "Cases Per 100,000")

```


```{r, echo=FALSE, include=FALSE}

#Performing similar analysis as the Choropleth processing above so we can do the same visualization as above with death rates, however, here I just want to include the deaths per capita in a formattable table as done previously for total counts by country.
 
death.state <- death.recent %>%
               filter(`Country/Region` == "US") %>%
               group_by(`Province/State`) %>%
               rename(state=`Province/State`,
                      total=count)

colnames(death.state)[colnames(death.state)=="total"] <- "deaths"


# Creating various subsets to join with the state data to create a single data table with percapita cases and deaths
death.tbl <- subset(death.state,select = c("deaths","state"))
state.tbl <- death.tbl %>% left_join(count.state, by=c("state"))
state.tbl$Death_Per_Capita <- round(state.tbl$deaths/state.tbl$pop_2015*100000,2)

# Find the top 10 states by number of cases
state.tbl <- state.tbl %>% arrange(desc(Confirmed_Per_Capita)) 
state.tbl.top <- top_n(ungroup(state.tbl), 15, Confirmed_Per_Capita)

#renaming the columns
colnames(state.tbl.top)[colnames(state.tbl.top)=="Confirmed_Per_Capita"] <- "Cases Per Cap"
colnames(state.tbl.top)[colnames(state.tbl.top)=="Death_Per_Capita"] <- "Deaths per Cap"
colnames(state.tbl.top)[colnames(state.tbl.top)=="state"] <- "States"
colnames(state.tbl.top)[colnames(state.tbl.top)=="deaths"] <- "Deaths"
colnames(state.tbl.top)[colnames(state.tbl.top)=="total"] <- "Cases"

state.tbl.top <- subset(state.tbl.top, select = c(States, Cases, `Cases Per Cap`, Deaths, `Deaths per Cap`))

state.tbl.top$Mortality <- percent((state.tbl.top$Deaths / state.tbl.top$Cases)) 

state.tbl.tbl <- formattable(state.tbl.top, align =c("l","r","r","r","r","r"),
                list(`SDOH` = formatter("span", style = ~ formattable::style(color = "grey",font.weight = "bold")),
                `Cases` = color_bar(customYel),
                `Cases Per Cap` = color_bar(customYel),
                `Deaths` = color_bar(customRed),
                `Deaths per Cap` = color_bar(customRed),
                `Mortality` = color_bar(customRed)))



```

## US Cases and Deaths Per Capita

Here are the top 15 States by case including per capita (per 100,000 population, 2015 census data) cases, deaths, and percent mortality. This will allow us to compare state by state efforts in containment. 


```{r, echo=FALSE, warning=FALSE}

formattable(state.tbl.tbl)

```


























